### 2018-12-19 code.tex
finished writing on Traffic mix

\chapter{Runway Usage}\label{cha:RWY_usage}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{graphics/fig_runway_usage_2017-10-04_to_2018-11-30.png}
    \caption[Runway usage at BIKF]{Overall runway usage at BIKF for a period of one year since October 2017. Almost half of all arrivals during that period (44,7\%) use RWY-19, followed by RWY-01 (25,7\%) and RWY-10 (17.6\%). The summer traffic favours RWY-19 with 52.8\%~(Figure~\ref{fig:runway_usage_summer}), while during the winter season some of those arrivals are diverted towards RWY-01 and RWY-10~(Figure~\ref{fig:runway_usage_summer}). RWY-28 is least used despite its fast-exit TWY B-1.}
    \label{fig:runway_usage}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]{graphics/fig_runway_usage_summer}
    \caption[Summer runway usage at BIKF]{The summer traffic favours RWY-19 (52,8\%), followed by RWY-01 (23,3\%), while RWY-10 and RWY-28 are with 11,6\% and 12,3\% respectively.}
    \label{fig:runway_usage_summer}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]{graphics/fig_runway_usage_winter}
    \caption[Winter runway usage at BIKF]{The winter traffic at BIKF is a bit more evenly distributed among runways bit still favours RWY-19 (38,8\%),followed by RWY-01 with 28,9\%, RWY-10 (20,8\%) and RWY-28 (11,4\%).}
    \label{fig:runway_usage_winter}
\end{figure}

\chapter{Arrival Runway Occupancy Times\label{cha:AROTs}}

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{lr|r|r|r|r|r|r|r|}
\cline{3-9}
\multicolumn{2}{l|}{RWY01}  & \multicolumn{7}{c|}{AROT {[}s{]}} \\ \hline
\multicolumn{1}{|l|}{Season} & count & \multicolumn{1}{l|}{mean} & \multicolumn{1}{l|}{std} & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{25\%} & \multicolumn{1}{l|}{50\%} & \multicolumn{1}{l|}{75\%} & \multicolumn{1}{l|}{max} \\ \hline
\multicolumn{1}{|l|}{Summer} & \multicolumn{1}{r|}{764} & 85 & 20 & 46 & 70 & 86 & 99 & 153 \\ \hline
\multicolumn{1}{|l|}{Winter} & \multicolumn{1}{r|}{263} & 100 & 17 & 55 & 87 & 98 & 111 & 156 \\ \hline
\end{tabular}%
}
\caption[AROTs RWY-01 pre fast exit by season]{AROTs for runway RWY-01 pre-fast-exit TWY A-1 by season}
\label{tab:season_AROT_stats_RWY01_pre_fast_exit}
\end{table}

% ----------------------------

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{lr|r|r|r|r|r|r|r|}
\cline{3-9}
\multicolumn{2}{l|}{RWY01, TWY A-1}  & \multicolumn{7}{c|}{AROT  {[}s{]}} \\ \hline
\multicolumn{1}{|l|}{Season} & count & \multicolumn{1}{l|}{mean} & \multicolumn{1}{l|}{std} & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{25\%} & \multicolumn{1}{l|}{50\%} & \multicolumn{1}{l|}{75\%} & \multicolumn{1}{l|}{max} \\ \hline
\multicolumn{1}{|l|}{Summer} & \multicolumn{1}{r|}{1879} & 84 & 22 & 46 & 64 & 86 & 98 & 152 \\ \hline
\multicolumn{1}{|l|}{Winter} & \multicolumn{1}{r|}{284} & 88 & 26 & 47 & 63 & 90 & 106 & 153 \\ \hline
\end{tabular}%
}
\caption[AROTs RWY-01 post fast exit by season]{AROTs for runway RWY-01 post-fast-exit TWY A-1 by season}
\label{tab:season_AROT_stats_RWY01_post_fast_exit}
\end{table}

% ----------------------------

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{lr|r|r|r|r|r|r|r|}
\cline{3-9}
\multicolumn{2}{l|}{RWY28} & \multicolumn{7}{c|}{AROT {[}s{]}} \\ \hline
\multicolumn{1}{|l|}{Season} & count & \multicolumn{1}{l|}{mean} & \multicolumn{1}{l|}{std} & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{25\%} & \multicolumn{1}{l|}{50\%} & \multicolumn{1}{l|}{75\%} & \multicolumn{1}{l|}{max} \\ \hline
\multicolumn{1}{|l|}{Summer} & \multicolumn{1}{r|}{553} & 99 & 12 & 46 & 91 & 98 & 106 & 153 \\ \hline
\multicolumn{1}{|l|}{Winter} & \multicolumn{1}{r|}{249} & 110 & 13 & 60 & 100 & 108 & 117 & 155 \\ \hline
\end{tabular}%
}
\caption[AROTs RWY-28 pre fast exit by season]{AROTs for runway RWY-28 pre-fast-exit TWY B-1 by season}
\label{tab:season_AROT_stats_RWY28_pre_fast_exit}
\end{table}

% ----------------------------

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{lr|r|r|r|r|r|r|r|}
\cline{3-9}
\multicolumn{2}{l|}{RWY28, TWY B-1} & \multicolumn{7}{c|}{AROT {[}s{]}} \\ \hline
\multicolumn{1}{|l|}{Season} & {count} & \multicolumn{1}{l|}{mean} & \multicolumn{1}{l|}{std} & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{25\%} & \multicolumn{1}{l|}{50\%} & \multicolumn{1}{l|}{75\%} & \multicolumn{1}{l|}{max} \\ \hline
\multicolumn{1}{|l|}{Summer} & 401 & 68 & 14 & 49 & 60 & 66 & 72 & 155 \\ \hline
\multicolumn{1}{|l|}{Winter} & 69 & 73 & 13 & 49 & 63 & 72 & 79 & 105 \\ \hline
\end{tabular}%
}
\caption[AROTs RWY-28 post fast exit by season]{AROTs for runway RWY-28 post-fast-exit TWY B-1 by season}
\label{tab:season_AROT_stats_RWY28_post_fast_exit}
\end{table}

% ----------------------------

% \chapter{Code}\label{cha:code}
% You can put code in your document using the listings package, which is
% loaded by default in \path{custom.tex}.  Be aware that the listings
% package does not put code in your document if you are in draft mode
% unless you set the \texttt{forcegraphics} option.

% There is an example java (Listing~\ref{src:Data_Bus.java}) and XML
% file (Listing~\ref{src:AndroidManifest.xml}).  Thanks to the
% \texttt{url} package, you can typeset OSX and unix paths like this:
% \path{/afs/rnd.ru.is/project/thesis-template}.  Windows paths:
% \path{C:\windows\temp\ }.  You can also typeset them using the menukey
% package, but it tends to delete the last separator and has other
% complications.\footnote{The menukey package has issues with biblatex,
%   read \path{custom.tex} for more information.}

% If you are trying to include multiple different languages, you should
% go read the documentation and set these up in \path{custom.tex}.  You
% will save yourself a lot of effort, especially if you have to fix
% anything.

% %I have put the source code in the \directory{src/} folder.
% \lstinputlisting[language=Java, firstline=1,
% lastline=40, caption={Data\_Bus.java: Setting up the class.},
% label={src:Data_Bus.java}]{src/Data_Bus.java}

% \lstinputlisting[language={[android]XML}, firstline=1, lastline=20,
% caption={AndroidManifest.xml: Configuration for the Android UI.},
% label={src:AndroidManifest.xml}]{src/AndroidManifest.xml}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "DEGREE-NAME-YEAR"
%%% End: